A Gentle Introduction into Structural Causal Models

### 2021-06-10 paper_scm.Rmd
sq

@@

@@

@@

@@

$N:= f_N(U_N)$\newline
$L:=f_L(U_L)$\newline
$E:= f_E(N,L,U_E)$\newline

$\{U_N\}$ - Numeric Ability

$\{U_L\}$ - Literary Ability

$\{U_E\}$ - Quality of the Student

$\{N\}$ - Math Test

$\{L\}$ - English-Test

$\{E\}$ - GPA

@@

 |Method          | Action | | Usage | 
|------------------|-------------|-------------------|
| Association $P(a|b)$               | Co-occurrence                           |(Un-)Supervised ML, BN, Reg.  
| Intervention $P(a|do(b),c)$       | Do-manipulation                       |CBN,MDP,RL    
| Counterfactual $P(a_b|a`,b`)$     | Hypotheticals            | SCM ,PO

@@

@@

@@

@@

>>>>

History

@@

$S:= f_S(U_S)$
$E:= f_E(U_E)$
$C:= f_C(S,E,U_C)$

$\{S\}$ - Smoking
$\{C\}$ - Cancer
$\{E\}$ - Eating

@@

Independence of Noise, Independence of Mechanism, Causal Sufficiency, and why time is ignored.
The question should be really, what are we commiting to when we use SCMs.

**Acyclicity**

Assume acyclic because conditional probabilities with variables causing itself very hard .

@@

Noise indepedent so noise smoking and noise eating should be independent.

@@

@@

 |Method          | Action | | Usage | 
|------------------|-------------|-------------------|
| Association $P(a|t)$               | Co-occurrence                           |(Un-)Supervised ML, BN, Reg.  
| Intervention $P(a|do(t),c)$       | Do-manipulation                       |CBN,MDP,RL    
| Counterfactual $P(a_t|a`,t`)$     | Hypotheticals            | SCM ,PO

@@

-> example probabilistic terms

$P(C|s)$

@@

$P(a|do(t),c)$

$do(t)$ -> replace function $T:=f_T(\pi)$ with different conditional probability (or constant) [@pearl2009causality].

@@

-> how to obtain scm

@@

largely ingore time

time in mechanical modeling crucial [@peters_elements_2017].

- Time in Social Sciences: Often Vague 
- Time in Physical Sciences: Mechanical via **Differential equations** 
- dependence on prior time point and change in time contribute to the value at time point

Initial Value: 
$$
\mathbf{x}\left(t_{0}\right)=\mathbf{x}_{0}
$$

Derivative of function x with respect to time t: 
$$
\frac{d \mathbf{x}}{d t}=f(\mathbf{x}), \mathbf{x} \in \mathbb{R}^{d}
$$

Value of Function at time t + dt: 
$$
\mathbf{x}(t+d t)=\mathbf{x}(t)+d t \cdot f(\mathbf{x}(t))
$$
